Prospective testing for the prevalence or transience of a shock effect over a set time horizon before it occurs

### 2022-06-11 prevalence_testing/postshocktest2022draft.tex
revision to numerical results

\begin{table}[H]
	\caption{Results for Conoco Philips ($n=3$), $\mbf{W}^* = (0.183, 0, 0.817)$ from scaled covariates} \label{coptable}
\begin{center}
  \begin{tabular}{cccccccccr}
     $H$  & Model &   $p_1$ & $\hat{p}_1$ & $|\hat{p}_1-p_1|$ & Vote (Correct?) & weighted Vote (Correct?)   \\
     \hline 
     \multirow{2}{*}{2} & $\mc{M}_0$  &   1 & 0.183  & 0.817 & 0 (Yes) & 1 (No)  \\      
      & $\mc{M}$  & 1 & 1 & $0$   & 0 (Yes) &0 (Yes)   \\[.2cm]
     \multirow{2}{*}{4} & $\mc{M}_0$ &   0 & 0.519  & $0.519$ & 0 (No) &1 (No)    \\
      &  $\mc{M}$ & 1& 1 & 0 &0 (Yes)  &0 (Yes)  \\[.2cm]
     \multirow{2}{*}{6} &  $\mc{M}_0$ &   0   & $0.817$ & 0.817 & 0 (No) &1 (No)   \\
      &  $\mc{M}$ & 1 & 1  & 0 &0 (Yes) & 0 (Yes) \\[.2cm]
     \multirow{2}{*}{8} & $\mc{M}_0$ & 0  & 0.817 & 0.817 & 0 (No) & 0 (No)\\
     & $\mc{M}$ & 1 & 1 & 0 & 0 (Yes) & 0 (Yes)
  \end{tabular}
\end{center}	
\end{table}

We now discuss the numerical results. Note that the synthetic weight vector $\mbf{W}^*$ for the three donors is $(0.183, 0, 0.817)$. It suggests that the donor associated with the shock on November 27, 2014 is the most similar time series relative to the time series of interest, and the one on March 15, 2008 is the second. As pointed out by \cite{lin2021minimizing} in Section 5.1(3), the shocks experienced by the first and  third donor are related to recessions, and oil supply shock, respectively. The results are intuitive in the sense that from a retrospective perspective of a data analyst, the shock experienced by the series $\{y_{1,t}\colon t = 1, \ldots, T_1\}$ is due to oil supply shock and recessions caused by COVID-19 pandemic.

Let $a_i^1$ and $a_i^2$ denote the AICs of $\mc{M}_0$ and $\mc{M}$ for $i$th time series, respectively. Based on \eqref{model-gen}, which describes $\mc{M}$ and the description of $\mc{M}$ in the beginning of Section \ref{sec::ne}, it may be inferred that $\mc{M}$ captures the post-shock dynamic change whereas $\mc{M}_0$ suggests a persistent post-shock additive location shift. Using the synthetic pairwise model selection procedures specified in Corollary \ref{coro2}, it turns out that $\sum_{i=2}^4 w_i^* I(a_i^1 > a_i^2)=1 >0.5$. Using this aggregated model selection approach, we prefer $\mc{M}$ over $\mc{M}_0$. \textcolor{red}{Add column for AIC results}

From Table \ref{coptable}, we can see that $p_1$ for $\mc{M}$ is not sensitive to the changes of $H$ and is uniformly 1. Under any reasonable significance level, based on the features of $\mc{M}$, we infer that post-shock dynamics do not differ from the pre-shock one and the shock is not persistent. In contrast, $p_1$ of $\mc{M}_0$ equals 1 for $H = 2$ whereas equals 0 for $H = 4, 6, 8$. It is possible that $\mc{M}_0$ does not fit the data well. In practice, a simple additive post-shock location-shift is rare and more complicated structures cannot be explained using this. To support this claim, we plotted $\{y_{1,t} \colon 1, \ldots, T_1\}$ in Figure \ref{copshocktransience}.

Two dashed vertical lines are plotted in Figure \ref{copshocktransience}, where the left one corresponds to $t = T_1^*$ and the right one  corresponds to $t = T_1^*+1$. As shown by \cite{lin2021minimizing}, $y_{1,t}$ experienced a substantial shock at  $t = T_1^*+1$ and it is evident from $y_{1, T_1^*}-y_{1, T_1^*+1}$ shown in Figure 1. We also graphed a benchmark horizontal dashed line for comparison, which corresponds to the stock price $y_{1, T_1^*}$. However, notice that even if $y_{1,t}$ first experienced a considerable drop at $t = T_1^*+1$,   it started to revert back and became closer to the benchmark horizontal line as $t$ increases for $t > T_1^*$. Note that this observation confirms our model selection and what $\mc{M}$ suggests for shock transience. It is also important to re-emphasize that  the test compares the forecasts from $t = T_1^* + H+1$ to $t=T_1$. In other words, the beginning dramatic drop as seen from Figure \ref{copshocktransience} around March is not considered in the forecast comparison. It is reasonable to our objective as we are interested in long-term persistence rather-than short-term substantial changes. It is   possible that the significance of $\mc{M}_0$ for $H = 2$ takes into account the short-term percussions of the substantial shock. It is because (1) for small $H$, the loss differential matrix incorporates some large loss differential elements corresponding to $t$ close to $T^*_1+1$ and (2) $\mc{M}_0$ can be very sensitive to substantial additive location shift around $t = T^*_1+1$. And this phenomenon disappears as $H$ increases.

@@

\textcolor{red}{We motivate a similar approach within our prevalence testing framework. The idea is to again separate out the financial and oil supply time series and add these series together which will create a new donor pool consisting of series which incorporate a dual shock whose behavior can be assessed with our methodology. As in \cite{lin2021minimizing} the oil supply shock donor pool consists of the single November 27, 2014 series; and the financial shock donor pool consists of the remaining two series. These added series which account for a potential dual shock take the following form where the subscript $o$ refers to the single November 27, 2014, the subscript $j$ indexes the other series, and $T_o = T_j$ and $T^*_o = T^*_j$}
\begin{equation} \label{eq:dualshock}
\begin{split}
  z_{j,t} &= y_{o,t} + y_{j,t} \\
  	&= \bar{\eta}_j + (\phi_o y_{o,t-1} +\theta_{o,t-1}'\x_{o,t-1} + \phi_j y_{j,t-1} +\theta_{j,t-1}'\x_{j,t-1})(1 - \bar{D}_t) \\
  	&\qquad + (\bar{\alpha}_j + \tilde{\phi}_o y_{o,t-1} + \tilde{\theta}_{o,t-1}'\x_{o,t-1} + \tilde{\phi}_j y_{j,t-1} + \tilde{\theta}_{j,t-1}'\x_{j,t-1})\bar{D}_t + \bar{\varepsilon}_t,
\end{split}
\end{equation}
where $\bar{\eta}_j = \eta_o + \eta_j$, $\bar{\alpha}_j = \alpha_o + \alpha_j$, $D_t = D_{0,t} = D_{j,t}$, and $\bar{\varepsilon}_t = \varepsilon_{0,t} + \varepsilon_{j,t} + D_t(\varepsilon_{\alpha,o} + \varepsilon_{\alpha,j})$. The model \eqref{eq:dualshock} is then a regression model with a heterogeneous error structure, the error distribution post-shock differs from the error distribution pre-shock. All model parameters are estimable in \eqref{eq:dualshock}. However, aggregation strategies are slightly different for this model since covariates are included for two separate series.

@@

@@

\begin{table}[H]
	\caption{Results for Month-on-Month Change in Log Nonfarm Payrolls (Seasonally Adjusted) ($n=3$), $\mbf{W}^* = (0, 0.278, 0.722)$ from scaled covariates}   \label{uemtable}
\begin{center}
  \begin{tabular}{cccccccccr}
     $H$  & Model &   $p_1$ & $\hat{p}_1$ & $|\hat{p}_1-p_1|$ & Vote (Correct?) & weighted Vote (Correct?)   \\
     \hline 
     \multirow{2}{*}{2} & $\mc{M}_0$  &   0.190 & 0.436  & 0.246 & 0 (Yes) & 0 (Yes)\\      
      & $\mc{M}$  & 0.520 & 0.556 & $0.036$   & 0 (Yes) &0 (Yes)   \\[.2cm]
     \multirow{2}{*}{3} & $\mc{M}_0$ &   0.140 & 0.413  & $0.273$ & 0 (Yes) & 0 (Yes)    \\
      &  $\mc{M}$ & 0.495 & 0.656 & 0.161 & 0 (Yes) & 0 (Yes)  \\[.2cm]
     \multirow{2}{*}{4} &  $\mc{M}_0$ &   0.275   & $0.464$ & 0.189 & 0 (Yes) & 0 (Yes)  \\
      &  $\mc{M}$ & 0.350 & 0.559  & 0.209 & 0 (Yes) & 0 (Yes) \\[.2cm]
     \multirow{2}{*}{5} & $\mc{M}_0$ & 0.380  & 0.275 & 0.105 & 0 (Yes) & 0 (Yes)\\
     & $\mc{M}$ & 0.665 & 0.763 & 0.098 & 0 (Yes) & 0 (Yes)
  \end{tabular}
\end{center}	
\end{table}

$\sum_{i=2}^4 w_i^* I(a_i^1 > a_i^2)=0.722$

>>>>

\begin{table}[H]
	\caption{Results for Conoco Philips ($n=3$), $\mbf{W}^* = (0.221, 0, 0.779)$ from scaled covariates} \label{coptable}
\begin{center}
  \begin{tabular}{cccccccccr}
     $H$  & Model &   $p_1$ & $\hat{p}_1$ & $|\hat{p}_1-p_1|$ & Vote (Correct?) & weighted Vote (Correct?)   \\
     \hline 
     \multirow{2}{*}{2} & $\mc{M}_0$  &   1 & 0.225  & 0.775 & 0 (Yes) & 1 (No)  \\      
      & $\mc{M}$  & 1 & 1 & $0$   & 0 (Yes) &0 (Yes)   \\[.2cm]
     \multirow{2}{*}{4} & $\mc{M}_0$ &   0 & 0.953  & $0.953$ & 0 (No) &0 (No)    \\
      &  $\mc{M}$ & 1& 1 & 0 &0 (Yes)  &0 (Yes)  \\[.2cm]
     \multirow{2}{*}{6} &  $\mc{M}_0$ &   0   & $1$ & 1 & 0 (No) &0 (No)   \\
      &  $\mc{M}$ & 1 & 1  & 0 &0 (Yes) & 0 (Yes) \\[.2cm]
     \multirow{2}{*}{8} & $\mc{M}_0$ & 0  & 1 & 1 & 0 (No) & 0 (No)\\
     & $\mc{M}$ & 1 & 1 & 0 & 0 (Yes) & 0 (Yes)
  \end{tabular}
\end{center}	
\end{table}

We now discuss the numerical results. Note that the synthetic weight vector $\mbf{W}^*$ for the three donors is $(0.221, 0, 0.779)$. It suggests that the donor associated with the shock on November 27, 2014 is the most similar time series relative to the time series of interest, and the one on March 15, 2008 is the second. As pointed out by \cite{lin2021minimizing} in Section 5.1(3), the shocks experienced by the first and  third donor are related to recessions, and oil supply shock, respectively. The results are intuitive in the sense that from a retrospective perspective of a data analyst, the shock experienced by the series $\{y_{1,t}\colon t = 1, \ldots, T_1\}$ is due to oil supply shock and recessions caused by COVID-19 pandemic.

Let $a_i^1$ and $a_i^2$ denote the AICs of $\mc{M}_0$ and $\mc{M}$ for $i$th time series, respectively. Based on \eqref{model-gen}, which describes $\mc{M}$ and the description of $\mc{M}$ in the beginning of Section \ref{sec::ne}, it may be inferred that $\mc{M}$ captures the post-shock dynamic change whereas $\mc{M}_0$ suggests a persistent post-shock additive location shift. Using the synthetic pairwise model selection procedures specified in Corollary \ref{coro2}, it turns out that $\sum_{i=2}^4 w_i^* I(a_i^1 > a_i^2)=1 >0.5$. Using this aggregated model selection approach, we prefer $\mc{M}$ over $\mc{M}_0$. \textcolor{red}{For reference, the AICs for $\mc{M}_0$ in the donors are $(297.410, 413.357, 295.174)$ whereas those for $\mc{M}$ in the donors are $(302.835, 409.544, 280.609)$. }

From Table \ref{coptable}, we can see that $p_1$ for $\mc{M}$ is not sensitive to the changes of $H$ and is uniformly 1. Under any reasonable significance level, based on the features of $\mc{M}$, we infer that post-shock dynamics do not differ from the pre-shock one and the shock is not persistent. In contrast, $p_1$ of $\mc{M}_0$ equals 1 for $H = 2$ whereas equals 0 for $H = 4, 6, 8$. It is possible that $\mc{M}_0$ does not fit the data well. In practice, a simple additive post-shock location-shift is rare and more complicated structures cannot be explained using this.

Two dashed vertical lines are plotted in Figure \ref{copshocktransience}, where the left one corresponds to $t = T_1^*$ and the right one  corresponds to $t = T_1^*+1$. As shown by \cite{lin2021minimizing}, $y_{1,t}$ experienced a substantial shock at  $t = T_1^*+1$ and it is evident from $y_{1, T_1^*}-y_{1, T_1^*+1}$ shown in Figure 1. We also graphed a benchmark horizontal dashed line for comparison, which corresponds to the stock price $y_{1, T_1^*}$. However, notice that even if $y_{1,t}$ first experienced a considerable drop at $t = T_1^*+1$,   it started to revert back and became closer to the benchmark horizontal line as $t$ increases for $t > T_1^*$. Note that this observation confirms our model selection and what $\mc{M}$ suggests for shock transience. It is also important to re-emphasize that  the test compares the forecasts from $t = T_1^* + H+1$ to $t=T_1$. In other words, the beginning dramatic drop as seen from Figure \ref{copshocktransience} around March is not considered in the forecast comparison. It is reasonable to our objective as we are interested in long-term persistence rather-than short-term substantial changes.

@@

\textcolor{red}{We motivate a similar approach within our prevalence testing framework. The idea is to again separate out the financial and oil supply time series and add these series together which will create a new donor pool consisting of series which incorporate a dual shock whose behavior can be assessed with our methodology. As in \cite{lin2021minimizing} the oil supply shock donor pool consists of the single November 27, 2014 series; and the financial shock donor pool consists of the remaining two series. These added series which account for a potential dual shock take the following form where the subscript $o$ refers to the single November 27, 2014, the subscript $j$ indexes the other series, and $T_o = T_j$ and $T^*_o = T^*_j$}
\begin{equation} \label{eq:dualshock}
\begin{split}
  z_{j,t} &= y_{o,t} + y_{j,t} \\
  	&= \bar{\eta}_j + (\phi z_{j,t-1} +\theta_{o,t-1}'\x_{o,t-1} +\theta_{j,t-1}'\x_{j,t-1})(1 - \bar{D}_t) \\
  	&\qquad + (\bar{\alpha}_j + \tilde{\phi} z_{j,t-1} + \tilde{\theta}_{o,t-1}'\x_{o,t-1}   + \tilde{\theta}_{j,t-1}'\x_{j,t-1})\bar{D}_t + \bar{\varepsilon}_t,
\end{split}
\end{equation}
where $\bar{\eta}_j = \eta_o + \eta_j$, $\bar{\alpha}_j = \alpha_o + \alpha_j$, $D_t = D_{0,t} = D_{j,t}$, and $\bar{\varepsilon}_t = \varepsilon_{0,t} + \varepsilon_{j,t} + D_t(\varepsilon_{\alpha,o} + \varepsilon_{\alpha,j})$. The model \eqref{eq:dualshock} is then a regression model with a heterogeneous error structure, the error distribution post-shock differs from the error distribution pre-shock. All model parameters are estimable in \eqref{eq:dualshock}. However, aggregation strategies are slightly different for this model since covariates are included for two separate series.

@@

\begin{table}[H]
	\caption{Results for Conoco Philips dual shock ($n=2$), $\mbf{W}^* = (0.756, 0.244)$ from scaled covariates} \label{coptable}
\begin{center}
  \begin{tabular}{cccccccccr}
     $H$   &  $\hat{p}_1$ &   Vote (Correct?) & weighted Vote (Correct?)   \\
     \hline 
     2  & 1    & 0 (Yes) &0 (Yes)   \\ 
     4  & 1 & 0(Yes)  &0 (Yes)  \\ 
     6   & 1  &0 (Yes) & 0 (Yes) \\ 
     8   & 1  & 0 (Yes) & 0 (Yes)
  \end{tabular}
\end{center}	
\end{table}

@@

\begin{table}[H]
	\caption{Results for Month-on-Month Change in Log Nonfarm Payrolls (Seasonally Adjusted) ($n=3$), $\mbf{W}^* = (0.196, 0.804, 0.000)$ from scaled covariates}   \label{uemtable}
\begin{center}
  \begin{tabular}{cccccccccr}
     $H$  & Model &   $p_1$ & $\hat{p}_1$ & $|\hat{p}_1-p_1|$ & Vote (Correct?) & weighted Vote (Correct?)   \\
     \hline 
     \multirow{2}{*}{2} & $\mc{M}_0$  &   0.450 & 0.511  & 0.061 & 0 (Yes) & 0 (Yes)\\      
      & $\mc{M}$  & 0.240 & 0.583 & $0.343$   & 0 (Yes) &0 (Yes)   \\[.2cm]
     \multirow{2}{*}{3} & $\mc{M}_0$ &   0.500 & 0.469  & $0.031$ & 0 (Yes) & 0 (Yes)    \\
      &  $\mc{M}$ & 1 & 0.576 & 0.424 & 0 (Yes) & 0 (Yes)  \\[.2cm]
     \multirow{2}{*}{4} &  $\mc{M}_0$ &   0.660   & $0.443$ & 0.217 & 0 (Yes) & 0 (Yes)  \\
      &  $\mc{M}$ & 1.000 & 0.494  & 0.506 & 0 (Yes) & 0 (Yes) \\[.2cm]
     \multirow{2}{*}{5} & $\mc{M}_0$ & 0.700  & 0.556 & 0.144 & 0 (Yes) & 0 (Yes)\\
     & $\mc{M}$ & 0.470 & 1.000 & 0.530 & 0 (Yes) & 0 (Yes)
  \end{tabular}
\end{center}	
\end{table}

$\sum_{i=2}^4 w_i^* I(a_i^1 > a_i^2)=0$.

AICs for $\mc{M}_0$ in the donor pool are $(-609.173, -657.146,-678.949)$

AICs for $\mc{M}$ in the donor pool are $(-601.743,-649.484,-684.147)$
